---
title: "rocmice"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

`rocmice` is an R package to work with ROC (Receiver Operating Curves) in 
multiply imputed data. Using multiple imputation in prediction modelling can 
be helpful when missingness patterns don't follow MCAR and complete case 
analysis, single imputation or similar methods could seriously bias the model. 
Using pooled ROC and AUC makes evaluation of such models easier.

# Installation

You can install the package with
```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("https://github.com/jonas-schropp/rocmice.git")
```

And then call it like any other package using 
```{r}
library(rocmice)
```

# Usage

## Use case 

To illustrate the use of `rocmice` we use an especially pathological simulated 
data set. To learn more how it was created and what makes ROC analysis on this 
data set tricky, call

```{r}
data(patho)
help("patho")
```


```{r}
library(plotROC)
library(ggplot2)
library(ggpubr)
```

## Plotting the ROC curve

First we pool the ROC for the imputed data sets using `pool_roc_rr`:

```{r}
pooled_roc <- pool_roc_rr(
    data = patho$patho_imp,
    score = "score",
    target = "outcome",
    unique_vals = NULL,
    fpr_vals = seq(from=0.001, to=0.999, by=0.001),
    backtransform = TRUE,
    ci.level = 0.95,
    corr = 0.5,
    verbose = TRUE
)
```

Then we plot the ROC for the complete data set and the complete cases from 
the data set with missing values using `plotROC::geom_roc` and add the pooled 
ROC created with `pool_roc_rr` using `ggplot::geom_step`:

```{r warning = FALSE, fig.height = 5, fig.width = 5}
ggplot() + 
  geom_roc(
    data = patho$patho_complete, 
    aes(m = score, d = outcome, color = 'complete data')
    ) +
  geom_roc(
    data = patho$patho_amp, 
    aes(m = score, d = outcome, color = 'complete cases')
    ) +
  geom_step(
    data = pooled_roc, 
    aes(fpr, roc, color = 'multiple imputation'), 
    size = 1
    ) +
  labs(x = "FPR", y = "TPR") +
  theme_minimal() +
  scale_color_manual(
    breaks = c(
      'complete data', 
      'complete cases', 
      'multiple imputation'
      ),
    values = c(
      'complete data'='blue', 
      'complete cases'='red', 
      'multiple imputation'='black'
      )
    ) +
  theme(
    legend.title = element_blank(),
    legend.position = "bottom"
    )
```

It is easy to see that the ROC created from multiply imputed data closely 
tracks the one on a complete data set without missing values, while the one 
on complete cases shows an overly optimistic bias.




